---
title: "Problem Set 3"
author: "Amy Tan"
format: html
editor: visual
---

## Github

Link: <https://github.com/amyqtan/gov51-ps3>

## Question 1.1

```{r}
library(tidyverse) 
library(glmnet) 
library(knitr) 
compas <- read_csv("data/raw/compas_clean.csv")
```

## Question 1.2

```{r}
nrow(compas) 
ncol(compas)

```

There are 7214 observations in the dataset and 17 variables in the dataset. The unit of observation appears to be a person.

## Question 1.3

```{r}
mean(compas$recidivism)
unique(compas$recidivism)
```

a\) The mean recidivism rate in the data is 0.45.

b\) The outcome appears to be binary

## Question 1.4

```{r}
summarytable <- data.frame(
  Variable = c("Recidivism", "Age", "Female", "Black", "Hispanic", 
               "Asian", "Other Race", "Native American", 
               "Priors Count", "Felony", "Juvenile Felony Count", 
               "Juvenile Misdemenour Count","Juvenile Other Count", "Has Juvenile Felony","Has Juvenile Misdemenour", "Age Under 25","Age 25 - 45"),
  Mean = c(mean(compas$recidivism),
           mean(compas$age),
           mean(compas$female),
           mean(compas$black),
           mean(compas$hispanic),
           mean(compas$asian),
           mean(compas$other_race),
           mean(compas$native_american),
           mean(compas$priors_count),
           mean(compas$felony),
           mean(compas$juv_fel_count),
           mean(compas$juv_misd_count),
           mean(compas$juv_other_count),
           mean(compas$has_juv_felony),
           mean(compas$has_juv_misd),
           mean(compas$age_under_25),
           mean(compas$age_25_45)),
  Std_Dev = c(sd(compas$recidivism),
           sd(compas$age),
           sd(compas$female),
           sd(compas$black),
           sd(compas$hispanic),
           sd(compas$asian),
           sd(compas$other_race),
           sd(compas$native_american),
           sd(compas$priors_count),
           sd(compas$felony),
           sd(compas$juv_fel_count),
           sd(compas$juv_misd_count),
           sd(compas$juv_other_count),
           sd(compas$has_juv_felony),
           sd(compas$has_juv_misd),
           sd(compas$age_under_25),
           sd(compas$age_25_45)))

summarytable
```

age has most, asian has least? not sure though bc relative??

## Question 1.5

```{r}

```

## Question 2.1

```{r}
set.seed(51)
n <- nrow(compas) 
train_idx <- sample(1:n, size = floor(0.8 * n)) 
train <- compas[train_idx, ] 
test <- compas[-train_idx, ]
```

There are 5771 observations in the train set and 1443 observations in the test set.

## Question 2.2

```{r}
recidivism_all <- lm(recidivism ~ age + female + black + hispanic + asian + other_race + native_american + priors_count + felony + juv_fel_count + juv_misd_count + juv_other_count + has_juv_felony + has_juv_misd + age_under_25 + age_25_45, data = train)

summary(recidivism_all)
```

a\) At the 5% level, age, female, black, priors_count, felony, juv_other_count, has_juv_misd, and age_25_45 are all statistically significant.

b\) For a one unit increase in priors_count, on average, recidivism increases by 0.0295, holding all other variables constant.

## Question 2.3

```{r}
prediction_train_ols <- predict(recidivism_all, newdata = train)
prediction_test_ols <- predict(recidivism_all, newdata = test)

actual_train <- train$recidivism
actual_test <- test$recidivism


train_RMSE_ols <- sqrt(mean((actual_train - prediction_train_ols)^2))
test_RMSE_ols <- sqrt(mean((actual_test - prediction_test_ols)^2))

train_RMSE_ols
test_RMSE_ols
test_RMSE_ols - train_RMSE_ols
```

a\) The in-sample RMSE is 0.461.

b\) The out-of-sample RMSE is 0.462.

c\) The gap is not large //appears not to be over fitting because of the small gap

## Question 2.4

```{r}
set.seed(51)
folds <- sample(rep(1:5, length.out = nrow(train)))

rmse_cv <- numeric(5) 
for (k in 1:5) { 
  train_k <- train[folds != k, ] 
  test_k <- train[folds == k, ] 
  model_k <- lm(recidivism ~ ., data = train_k) 
  pred_k <- predict(model_k, newdata = test_k) 
  rmse_cv[k] <- sqrt(mean((test_k$recidivism- pred_k)^2)) 
  } 

mean(rmse_cv)
```

## Question 3.1

```{r}
f_full <- recidivism ~ (.)^2 + 
  I(age^2) + I(priors_count^2) + 
  I(juv_fel_count^2) + I(juv_misd_count^2) + 
  I(juv_other_count^2)

X_train <- model.matrix(f_full, data = train)[,-1] 
X_test <- model.matrix(f_full, data = test)[,-1] 
y_train <- train$recidivism 
y_test <- test$recidivism

ncol(X_train)
nrow(X_train) 
nrow(X_train) / ncol(X_train)
```

a\) The expanded model has 141 predictors.

b\) The ratio of training observations to predictors is 40.9.

## Question 3.2

```{r}
ols_full <- lm(f_full, data = train)

prediction_train_ols_full <- predict(ols_full, newdata = train)
prediction_test_ols_full <- predict(ols_full, newdata = test)


train_RMSE_ols_full <- sqrt(mean((actual_train - prediction_train_ols_full)^2))
test_RMSE_ols_full <- sqrt(mean((actual_test - prediction_test_ols_full)^2))

train_RMSE_ols_full
test_RMSE_ols_full
test_RMSE_ols_full - train_RMSE_ols_full
```

a\) The in-sample RMSE is 0.4485.

b\) The out-of-sample RMSE is 0.4890.

c\) the gap is larger bc overfitting, fitting on noise

## Question 3.3

The in-sample RMSE is lower, showing how the new model fits better in-sample since the new model with 180 interaction terms is likely fitting on noise within the sample. However, the out-of-sample prediction worsens because there is overfitting occurring, where the model is fitting on high variance within the train sample, creating a greater difference in RMSE between the train and test sample.

## Question 4.1

```{r}
set.seed(51) 
cv_ridge <- cv.glmnet(X_train, y_train, alpha = 0, nfolds = 10)
cv_ridge$lambda.min
plot(cv_ridge)
```

a\) The optimal lambda value selected by cross-validation is 0.0138.

b\) ?

## Question 4.2

```{r}
prediction_test_ridge <- predict(cv_ridge, newx = X_test, s= "lambda.min")

test_ridge_RMSE <- sqrt(mean((actual_test - prediction_test_ridge)^2))
test_ridge_RMSE
```

## Question 4.3

```{r}
coef(cv_ridge, s = "lambda.min")
ridge_coefs <- coef(cv_ridge, s = "lambda.min")
ridge_nonzero <- rownames(ridge_coefs)[ridge_coefs[, 1] != 0]
ridge_nonzero <- ridge_nonzero[ridge_nonzero != "(Intercept)"]
ridge_nonzero
```

did not eliminate any variables bc that's what ridge does

RIdge did not set any coefficients to zero.

## Question 5.1

```{r}
set.seed(51) 
cv_lasso <- cv.glmnet(X_train, y_train, alpha = 1, nfolds = 10)
cv_lasso$lambda.min
plot(cv_lasso)
```

a\) The optimal lambda is 0.007694.

## Question 5.2

```{r}
lasso_coefs <- coef(cv_lasso, s = "lambda.min")
lasso_nonzero <- rownames(lasso_coefs)[lasso_coefs[, 1] != 0]
lasso_nonzero <- lasso_nonzero[lasso_nonzero != "(Intercept)"]

lasso_coefs
lasso_nonzero
```

a\) 27 variables survived and 114 were eliminated.

b\) Some of the interaction terms make sense, but some do not. For example,

age:age_25_45 does not make sense but black:juv_fel_count makes sense (being black makes juvenile felony count matter more.

c\) you're not basing in theory but rather numerical?

## Question 5.3

```{r}
prediction_test_lasso <- predict(cv_lasso, newx = X_test, s= "lambda.min")


test_lasso_RMSE <- sqrt(mean((actual_test - prediction_test_lasso)^2))
test_lasso_RMSE
```

The out of sample RMSE for lasso is 0.4593. how does it compare?

## Question 6.1

```{r}
set.seed(51) 
cv_elasticnet <- cv.glmnet(X_train, y_train, alpha = 0.5, nfolds = 10)

prediction_test_elasticnet <- predict(cv_elasticnet, newx = X_test, s= "lambda.min")


test_RMSE_elasticnet <- sqrt(mean((actual_test - prediction_test_elasticnet)^2))
test_RMSE_elasticnet

elasticnet_coefs <- coef(cv_elasticnet, s = "lambda.min")
elasticnet_nonzero <- rownames(elasticnet_coefs)[elasticnet_coefs[, 1] != 0]
elasticnet_nonzero <- elasticnet_nonzero[elasticnet_nonzero != "(Intercept)"]
elasticnet_nonzero
```

The out-of-sample RMSE is 0.4583 and it retains 43 variables.

## Question 6.2

```{r}


data.frame(
  Variable = c("OLS (main effects)", "OLS (kitchen sink)", "Ridge", "LASSO", "Elastic Net"),
  Predictors = c(ncol(compas) - 1,
           196,
           141,
           27,
           43),
  Out_of_Sample_RMSE = c(test_RMSE_ols,
    test_RMSE_ols_full,
    test_ridge_RMSE,
    test_lasso_RMSE,
    test_RMSE_elasticnet))
```

## Question 6.3

```{r}

```

## Question 6.4

```{r}

```

## Question 6.5

```{r}

```
