---
title: "Problem Set 3"
author: "Amy Tan"
format: html
editor: visual
---

## Github

link

## Question 1.1

```{r}
library(tidyverse) 
library(glmnet) 
library(knitr) 
compas <- read_csv("compas_clean.csv")
```

## Question 1.2

```{r}
nrow(compas) 
ncol(compas)

```

There are 7214 observations in the dataset and 17 variables in the dataset. The unit of observation appears to be a person.

## Question 1.3

```{r}
mean(compas$recidivism)
unique(compas$recidivism)
```

a\) The mean recidivism rate in the data is 0.45.

b\) The outcome appears to be binary

## Question 1.4

```{r}
summarytable <- data.frame(
  Variable = c("Recidivism", "Age", "Female", "Black", "Hispanic", 
               "Asian", "Other Race", "Native American", 
               "Priors Count", "Felony", "Juvenile Felony Count", 
               "Juvenile Misdemenour Count","Juvenile Other Count", "Has Juvenile Felony","Has Juvenile Misdemenour", "Age Under 25","Age 25 - 45"),
  Mean = c(mean(compas$recidivism),
           mean(compas$age),
           mean(compas$female),
           mean(compas$black),
           mean(compas$hispanic),
           mean(compas$asian),
           mean(compas$other_race),
           mean(compas$native_american),
           mean(compas$priors_count),
           mean(compas$felony),
           mean(compas$juv_fel_count),
           mean(compas$juv_misd_count),
           mean(compas$juv_other_count),
           mean(compas$has_juv_felony),
           mean(compas$has_juv_misd),
           mean(compas$age_under_25),
           mean(compas$age_25_45)),
  Std_Dev = c(sd(compas$recidivism),
           sd(compas$age),
           sd(compas$female),
           sd(compas$black),
           sd(compas$hispanic),
           sd(compas$asian),
           sd(compas$other_race),
           sd(compas$native_american),
           sd(compas$priors_count),
           sd(compas$felony),
           sd(compas$juv_fel_count),
           sd(compas$juv_misd_count),
           sd(compas$juv_other_count),
           sd(compas$has_juv_felony),
           sd(compas$has_juv_misd),
           sd(compas$age_under_25),
           sd(compas$age_25_45)))

summarytable
```

age has most, asian has least? not sure though bc relative??

## Question 1.5

```{r}

```

## Question 2.1

```{r}
set.seed(51)
n <- nrow(compas) 
train_idx <- sample(1:n, size = floor(0.8 * n)) 
train <- compas[train_idx, ] 
test <- compas[-train_idx, ]
```

There are 5771 observations in the train set and 1443 observations in the test set.

## Question 2.2

```{r}
recidivism_all <- lm(recidivism ~ age + female + black + hispanic + asian + other_race + native_american + priors_count + felony + juv_fel_count + juv_misd_count + juv_other_count + has_juv_felony + has_juv_misd + age_under_25 + age_25_45, data = train)

summary(recidivism_all)
```

a\) At the 5% level, age, female, black, priors_count, felony, juv_other_count, has_juv_misd, and age_25_45 are all statistically significant.

b\) For a one unit increase in priors_count, on average, recidivism increases by 0.0295, holding all other variables constant.

## Question 2.3

```{r}
prediction_train <- predict(recidivism_all, newdata = train)
prediction_test <- predict(recidivism_all, newdata = test)

actual_train <- train$recidivism
actual_test <- test$recidivism


train_RMSE <- sqrt(mean((prediction_train - actual_train)^2))
test_RMSE <- sqrt(mean((prediction_test - actual_test)^2))

train_RMSE
test_RMSE
```

a\) The in-sample RMSE is 0.461.

b\) The out-of-sample RMSE is 0.462.

c\) The gap is not large //appears not to be over fitting because of the small gap

## Question 2.4

```{r}

```
